Deep Neural Network for Nuclei Segmentation in Histopathology images

### 2016-11-04 Paper.tex
first update

% Template for ISBI-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

@@

Talk about related work.
Recent work in deep learning for segmentation, histopathology analysis helped by images. Maybe not too much as the paper is small..
\section{DataSet}
\label{sec:dataset}
Creating of the data set, some features about it. ITK-snap, some statistics about it.
\section{Methodology}
\label{sec:method}
About different architectures described. Validation scheme.

\section{Result}
\label{sec:result}
table of values
\section{Discussion}
\label{sec:discussion}
What's next?

@@

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{image1}}
%  \vspace{2.0cm}
  \centerline{(a) Result 1}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{image2}}
%  \vspace{1.5cm}
  \centerline{(b) Results 3}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{image3}}
%  \vspace{1.5cm}
  \centerline{(c) Result 4}\medskip
\end{minipage}
%
\caption{Example of placing a figure with experimental results.}
\label{fig:res}
%
\end{figure}

>>>>

% Template for ISBI-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{amsfonts}
% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

@@

Talk about related work.
Recent work in deep learning for segmentation, histopathology analysis helped by images. Maybe not too much as the paper is small..
\section{DataSet}
\label{sec:dataset}

%% Contribution as a dataset, small introduction
One of the main contributions of this paper is the now public available 
nuclei detection dataset within HE stainned histopathology images which 
can be found at www.cbio.ensmp.fr:/pnaylor/download/dataset.zip (not 
a true address). This annotated dataset provides images clustered by 
patients. Each patient has at least 3 annotated $512 \times 512$ HE 
stainned histopathology images with their associated ground truth. Each 
ground truth image is a $512 \times 512$ where each pixel value above 
$0$ is considered as a pixel belonging to a nuclei. The differences in 
values of these pixels denote different nucleus, such an annotation can 
be used in several processing step which does or doesn't take into 
account clutered nuclei. See figure \ref{fig:annotation} for an example of 
three annotated images. This annotation was conducted via the help of 
the software ITK-snap and were annotated by the authors of the paper.

%% Data set creation procedure
These patients were randomly picked from an unpublished study on
tripple negative breast cancer. For each of these patients we had access
to their biopsy sample as a whole slide image (WSI). WSI enables a
medical practitionner to digitize the huge amount of information that
can be found in glass slides. WSI can be up to 60 GB big
uncompressed and can't be stored in RAM on a standard computer. Given
the WSI of a patient, we randomly cropped $512 \times 512$ samples 
from the WSI. %% explain that we choose the tissue area? 
3 to 7 images were choosen from the randomly sampled images to try and give the most diversed dataset among these patients. Once the samples were choosen, we fully annotated each nuclei via the software ITK-snap and touching nuclei are differentiate via a different annotation value.

%%Data set description 
We can describe the dataset provided.

\section{Methodology}
\label{sec:method}
%% Segmentation task
Let A be the space of RGB images, A can typically be 
$\mathbb{R}^{n \times p \times 3}$ and let B the space B the space of annotation 
images, in our case $\{0,1\}^{n \times p}$. We have a set of 
$(A_l,B_l)_{l \in [1, N]}$ for a supervised learning approach. Our goal is 
a prediction task 
named as semantic segmentation, we wish to maximize the prediction of 
an unseen element belonging to B given an new element in A. We maximize 
thus prediction by modelizing our prediction function as the softmax output of a deep neural 
network. We find the model parameters by minimizing a log loss function 
defined as:
 $\frac{1}{\sum_{i,j}w_{ij}} \sum_{i,j} \sum_k w_{i,j} t_{i,j,k} \log (\widehat{p_{i,j,k}})$
, where $k$ designates a certain label, $w_{i,j}$ is 
certain weight given to pixel $i,j$, $t_{i,j,k}$ is equal to 1 if pixel $i,j$ is 
of class $k$ and $\widehat{p_{i,j,k}}$ designates the estimated probability 
of pixel $i,j$ of being $k$ via the softmax output of the neural network.
We minimize the loss function via stochastic gradient descent.

%% Validation scheme
We have our training set $(A_l,B_l)_{l \in [1, N]}$ where N is equal to 
34, how ever each element $A_l$ belongs to a certain patients and 
several elements $A_l$ can belong to the same patient. As we are 
dealing with histopathology images, it is know that samples can widely 
vary from one patient to the other. We thus validate our model by a 
leave one patient out scheme. Our validation scheme is as followed, for a 
given set of hype parameters, we 
train our model on every patient except one that is used for validation. 
Our final score is averaged over all patients.
Several metrics assess the quality of the model, the accuracy (Acc), the 
F1 score and a performance (Perf) score which is the mean between the 
true positive rates and the true negative rates.

%% Data augmentation and Hyper parameter training: Training scheme
%% Check number of images and number of transformations
To train our models, as the number of available annotated is scarse, we 
used a great number of transformation for the data augmentation. From 
a original size of 33 annotated images, adding flips, rotations, bluriness 
and random elastic deformations enabled us to have more then 
$400 000$ training images per patient. 
We also try out several hyperparameter configurations: the learning rate and 
momentum for the stochastic gradient descent, the weight decay value.  
In practice, we found that hyperparameters tuning didn't influence the 
scores much, the exepction being the learning rate. If the learning rate 
was not of the right magnitude the given network did not seem to learn. 
We also experienced with different 
initialization value and if possible, we also considered pretrained layers. 
Using pretrained layers made learning more efficient and made score 
values more robust.

\section{Different architectures/Results}
\label{sec:method}
We experience with 4 know arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned.
\section{Result}
\label{sec:result}
table of values
\section{Discussion}
\label{sec:discussion}
What's next?

@@

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{RGB_1}}
%  \vspace{2.0cm}
  \centerline{Patient 498959}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}
